---
title: "Developing models to predict cytotoxicity of metal nanoparticles"
author: "TXT"
date: "12/1/2021"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup R packages for model development

After installing R and RStudio, open RStudio and run the following codes to install and load necessary packages for developing models and making web applications::

```{r}
packages <- c("svDialogs", "data.table", "data.table", "openxlsx", "mlbench", "caret", 
              "tools", "DT", "magrittr", "ggplot2", "nortest", "tseries", "stringr", 
              "RcmdrMisc", "lmtest", "dplyr", "randomForest", "shiny", "shinydashboard")

install.packages(setdiff(packages, rownames(installed.packages())))

library(svDialogs) 
library(data.table)
library(openxlsx)
library(mlbench)
library(caret)
library(tools)
library(DT)
library(ggplot2)
library(car)
library(nortest)
library(tseries)
library(RcmdrMisc)
library(lmtest)
library(dplyr)
library(randomForest)
```

Developing models for predicting cytotoxicity of metal nanoparticles

Choose excel file containing dataset and read the dataset:
```{r}
DataMetal <- read.xlsx("MetalESN.xlsx", sheet = 1, startRow = 1, colNames = TRUE,
                       rowNames = FALSE, detectDates = FALSE, skipEmptyRows = TRUE,
                       skipEmptyCols = TRUE, rows = NULL, cols = NULL, 
                       check.names = FALSE, namedRegion = NULL, 
                       na.strings = "NA", fillMergedCells = FALSE)

DataMetal <- select(DataMetal, c("Toxicity",
                                 "Dose",
                                 "Assay",
                                 "Time",
                                 "Species",
                                 "Cancer",
                                 "Cell_Tissue",
                                 "Cell_line",
                                 "SSA",
                                 "Zeta",
                                 "HSize",
                                 "CoreSize",
                                 "Coating",
                                 "Shape",
                                 "Metal"))
```

Split data into training and test set (70/30)
```{r}
set.seed(1991)
split_size <- floor(0.70 * nrow(DataMetal))
in_rows <- sample(c(1:nrow(DataMetal)), size = split_size, replace = FALSE)
train <- DataMetal[in_rows, ]
test <- DataMetal[-in_rows, ]
```

Train Random Forest model:
```{r}
train.control <- trainControl(method = "repeatedcv", number = 10, repeats = 5)
RFmodel <- train(Toxicity ~ ., data = train, method = "rf", ntree = 100, trControl = train.control)
print(RFmodel)
```

Use RFmodel to predict test set:
```{r}
predictions <- RFmodel %>% predict(test); predictions_train <- RFmodel %>% predict(train)

```

Get confusion matrix and performance of model:
```{r}
CMatrix <- confusionMatrix(predictions, as.factor(test$Toxicity))
Performance <- data.frame(Parameter = row.names(as.data.frame(CMatrix$byClass)),
                          Value = as.data.frame(CMatrix$byClass))
colnames(Performance) <- c("Parameters", "Values")
CMatrix
Performance
```

Save data and models for later use in web application:
```{r}
save(RFmodel, file = "RFmodel.RData")
```
